NorduGrid Storage -- Overview and Design of a Service-Oriented Storage System

### 2008-12-06 doc/storage_system/ARCStorage_src/ARCStorage.tex
Final version before submitting

\begin{document}
\pagestyle{headings} 
\title{Service-Oriented ARC Storage; \\An Overview and Design}
%% \author{  Jon K. Nilsen$^{2,3}$ \and Salman Toor$^{1}$
%%   \and Zsombor Nagy$^{5}$ \and Bjarte Mohn$^{4}$  \\ $^{1}$Scientific
%%   Computing Division, IT Department, Uppsala University, Sweden
%%   \\ $^{2}$Center for Information Technology, University of Oslo, Norway
%%   \\ $^{3}$Department of Physics,  University of Oslo, Norway
%%   \\  $^{4}$Department of Physics and Astronomy, Uppsala University,
%%   Sweden \\ $^{5}$National Information Infrastructure Development
%%   Institute, Hungary}
\author{  Jon K. Nilsen$^{\dagger,\ddagger}$ \and Salman Toor$^{*}$
  \and Zsombor Nagy$^{\mathparagraph}$ \and Bjarte Mohn$^{\mathsection}$ 
\\$^{*}$Dept. of Information Technology, Div. of Scientific Computing, Uppsala
University,\\ Box~337, SE-75105~Uppsala, Sweden
 \\ $^{\dagger}$University of Oslo, Center for Information Technology,\\
P.~O.~Box~1059, Blindern, N-0316~Oslo, Norway
  \\ $^{\ddagger}$University of Oslo, Dept. of Physics, P.~O.~Box~1048,
Blindern, N-0316~Oslo, Norway
  \\  $^{\mathsection}$Dept. of Physics and Astronomy, Div. of Nuclear and
Particle Physics,\\ Uppsala University, Box~535, SE-75121~Uppsala, Sweden \\
$^{\mathparagraph}$Institute of National Information and Infrastructure
Development\\ NIIF/HUNGARNET, Victor~Hugo~18-22, H-1132~Budapest, Hungary\\ \\ 
j.k.nilsen@usit.uio.no \hspace{5 mm}
  salman.toor@it.uu.se \hspace{5 mm} zsombor@niif.hu \\  \hspace{7 mm}
  bjarte.mohn@fysast.uu.se}

@@

\maketitle             
%% \begin{abstract}
%% The KnowARC project will deliver the next generation Advanced Resource
%% Connector Fall  2009. A major component of this
%% service-oriented Grid middleware will be its storage system, the
%% ARC Storage, of which a
%% prototype is due Spring 2009. We will in this paper present an
%% overview of the storage system, its design and architecture. We will
%% also present some first, proof-of-concept test results, deploying the
%% the storage with a geographically distribution among three different countries.
%% ~\\~\\~
%% \end{abstract}
\begin{abstract}
There is an ever increasing need to utilize geographically distributed
hardware resources, both in terms of CPU and in terms of storage. The
service-oriented architecture provides a natural framework for
managing these resources. The next generation Advanced Resource
Connector (ARC) is a service-oriented Grid solution that will provide
the middleware to represent distributed resources in one simple
framework. In this paper, we will present an overview of the ARC
storage system, itself a set of services providing a self-healing,
flexible Grid storage solution. We will also present some first
proof-of-concept test results from a deployment of the storage system
distributed across three different countries. 
~\\~\\~
\end{abstract}
%-----------------------------------------------------------
\section{Introduction}
\label{Introduction}

@@

In the advent of the next generation of the Advanced Resource
Connector (ARC) Grid middleware~\cite{arc} (new release due during
fall 2009), we present the ARC storage
system~\cite{ARCStoragedesigndoc}. This distributed storage system is
designed to provide an easy to use, flexible and scalable system that
can offer native storage and at the same time provide access to
third-party solutions like dCache~\cite{dCache,DSSWithdCache} by using
the same, uniform interface. Being part of ARC, the storage system is
based on a service-oriented architecture, in which each major
component of the system runs as a separate service within the ARC
Hosting Environment Daemon (HED)~\cite{HEDdesigndoc}. The HED service
container gives the capability of flexible replacement of the
components as well as the possibility to introduce modifications in
the future. %% The HED service container gives
%% the capability of flexible replacement of the components as well as
%% the future modification/enhancement of the system.

This paper is organized as follows. After describing
related work in Section \ref{Related Work}, we give a bird's eye
view of the next-generation Advanced Resource Connector (ARC) in
Section \ref{The Advance Resource Connector}. An overview of the
ARC Storage is given in Section \ref{The ARC Storage System}, while
the architecture of the storage system is elaborated in Section
\ref{Architecture of the ARC Storage}. In Section \ref{Testing and
  Discussion} we give some early, proof-of-concept results, before 
concluding in Section \ref{Conclusion and Future Work}.

@@

a joint effort between DESY \cite{dCachesite}, Fermi National Accelerator Laboratory \cite{FermiLab}, the  Nordic Data

@@

The ARC Storage, being a light-weight and flexible

@@

high bandwidth environments. In contrast, ARC Storage is optimized

@@

To deliver the non-trivial quality of services required by the Grid,
there are a number of services running inside the HED. For example, Grid
job execution and management is handled by the A-REX service 
\cite{AREXdesigndoc}, policy
decisions are taken by the Charon service, the ISIS service is
responsible for information indexing, batch job submission is
handled by the Sched service, etc. In the later sections our discussion
will focus on the architecture and the design of another major set
of services, i.e., the ARC storage.

\section{The ARC Storage System}
\label{The ARC Storage System}

The ARC Storage consists of a set of SOAP based services residing
within HED. Together, the services provide a self-healing, reliable, robust,
scalable, resilient and consistent data storage system. Data is
managed in a hierarchical global namespace with files and
subcollections grouped into collections\footnote{A concept very
  similar to files and directories in most common file systems.}.  A
dedicated root collection serves as a reference point for accessing the namespace. The hierarchy can then be
referenced using Logical Names. The global namespace is accessed
in the same manner as in local filesystems.

\begin{figure}
  \begin{center}
    \includegraphics[width=0.8\columnwidth]{arc1-storage-services-with-gateway.pdf}
  \end{center}
  \caption{Schematic of the ARC Storage architecture. The figure shows
    the main services of ARC Storage and the communication channels. The B stands for Bartender, the S Shepherd, the L Librarian, the G Gateway and the A-H stands for A-Hash. The straight lines denote the communication between services.}
  \label{fig:arc1_storage_service}
\end{figure}

Being based on a service-oriented architecture, the ARC
Storage consists of a set of services as shown in Fig.
\ref{fig:arc1_storage_service}. The services are as follows: The Bartender (B)
provides the high-level interface to the user; the Librarian (L)
handles the entire storage namespace, using the A-Hash (A-H) as a metadatabase; the
Gateway (G) provides access to third-party storage systems; and
the Shepherd (S) is the frontend for the physical storage node. 
See Section \ref{Architecture of the ARC Storage} for a detailed
discussion of the different services.   
The services communicate with each other through the
Message Chain Components in HED. The communication channels are
depicted by straight lines in Fig. \ref{fig:arc1_storage_service}.

The system supports file transfer through several transfer protocols,
with client side tools that hide various technical
details such as protocol specification, port numbers
and so on. To provide fault-tolerance, the system implements
automatic file replication, where
the replicas of a file\footnote{In this paper, a file
denotes an entry in the global namespace, while replica denotes the
physical file stored on a storage node.} are always stored on separate
storage nodes\footnote{A
  storage node is a server with a Shepherd, a Shepherd backend
and some storage service.}. To ensure a resilient, self-healing system, the Shepherd
regularly sends heartbeats to one of the Librarians. If the Shepherd fails
to send a heartbeat, one of the Librarians will automatically initiate
re-replication of the replicas between the other Shepherds.

In the default implementation, the services will provide a
full-featured and consistent data storage system using files as an
atomic unit. The scope of the ARC storage system in the forseeable
future is restricted to providing support for file-based data services.
%% there are currently no plans for intrinsic support for database
%% access.

The ARC Storage supports third-party storage services in two
ways. Using the Gateway service, files already stored in some
third-party storage can be accessed using the client tools provided by
the ARC Storage. To make use of third-party services for storing new
files, with the replication and fault-tolerance offered by the ARC
Storage, third-party storage elements can also be used as backends for the
Shepherd service.

External grid middleware components can
access the ARC storage system using ARC data service
interfaces directly. ARC will also provide interface
components that communicate via standard protocols such as SRM, which
will provide a single access point to the whole system.

\section{Architecture of the ARC Storage}
\label{Architecture of the ARC Storage}

@@

Each object in the ARC Storage has a globally
unique ID. A collection contains files and other collections,
and each of these entries has a name unique within the collection very
much like entries in a standard directory on a local filesystem. Besides
files and collections, the ARC Storage has a third type of entry
called mount-points, which are references to the third-party storages
within the global namespace.

Replicas in a distributed storage system can have different states;
they can be broken, deleted, partially uploaded, etc. In the
ARC Storage, all replicas have assigned a state, some of which are
`\textbf{alive}' (if the replica passed the checksum test, and the
Shepherd reports that the storage node is healthy), `\textbf{invalid}'
(if the replica has a wrong checksum, or the Shepherd claims it has no
such replica) and `\textbf{creating}' (if the replica is in
the state of uploading).

@@

The Shepherd services run as front-ends on storage nodes. A Shepherd
service reports to a Librarian about the 
node's health state in terms of replicas. While the Bartender initiates
file transfers, the actual transfers go directly between the Shepherd
and the clients.

@@

A-Hash is a hash table for consistently storing data in property-value
pairs. All metadata about files, collections, mount point, Shepherd's
health status, and so forth, is stored in the A-Hash. As the A-Hash is
the service storing the entire state of the storage system, it is
absolutely crucial for the ARC Storage that the A-Hash is
consistent. The distribution and replication of this service is
therefore both necessary and challenging.

\begin{figure}
\centering
\includegraphics[width=1.0\columnwidth]{arc1-storage-downloading.pdf}
\caption{Schematic showing the scenario of file download from the ARC Storage.}
\label{arc1-storage-downloading}
\end{figure}

\subsection{Features}
\subsubsection{Heartbeat monitoring}
\label{Heartbeat and Replication}
In the proposed architecture, each Shepherd periodically sends
heartbeats to a Librarian with information about replicas
whose state changed since the last heartbeat, and the corresponding
GUIDs. These heartbeats are then stored in the A-Hash, making them
visible to all the Librarians in the system. If any of the Librarians
notices that a Shepherd is late with its heartbeat, it will invalidate
all the replicas in that Shepherd.

@@

\subsubsection{Deletion}
\label{Deletion}
In a replicating storage system, there are several possible solutions
for file deletion, since both the replicas and the metadata need to be
removed. In our solution, we use the process of \textit{lazy
  deletion}, \cite{LazyDeletion}. When a client requests that a file
should be deleted (and the user has the proper permissions) the
Bartender instructs the Librarian to remove the file's GUID from the
A-Hash. When the Shepherd, periodically checking all its replicas,
discovers that a replica has no file (and hence, no Location), it will
automatically delete the replica.

@@

Due to the unpredictable nature of the Grid environment, it is essential to
have some degree of  automatic recovery system in case of unexpected
failures. Fault tolerant behavior is required both at the level of
metadata and on the level of physical storage. While the work on fault-tolerant metadata is still in
progress, the following two
scenarios will explain the currently available recovery mechanism for
physical storage: 
\begin{itemize}
\item In the case of a file having an invalid checksum, the Shepherd
immediately informs the Librarian and the Librarian changes the state of the given replica to
`\textbf{invalid}'. To recover its replica, the Shepherd contacts a
Bartender and asks for another replica of the file. The Bartender chooses a valid
replica, initiates a file transfer from a Shepherd having the replica,
and returns the TURL to the Shepherd with the invalid replica. When the
Shepherd has recieved the replica and compared the checksum, it notifies the
Librarian that the replica is alive again.  
\item In the case of a Shepherd going offline, a Librarian will,
  as mentioned earlier, notice the lack of heartbeats and invalidate all the
  replicas, initiating new replication for all the files stored in
  this storage node. However, if the Shepherd
  again comes online, there will evidently be more replicas than
  needed. The first Shepherd to notice this will set its replica's
  state to `\textbf{thirdwheel}', i.e. obsolete. At the next occasion,
  the Shepherd will remove the replica, if and only if it has the only
  `\textbf{thirdwheel}' replica of this file. If there are more
  replicas with this state, all replicas will be set back to
  `\textbf{alive}' and the process is repeated. This scenario will be discussed further in Section
  \ref{Testing and Discussion}.
\end{itemize}
\subsubsection{Client tools }
\label{Client extension using FUSE module}

Being the only part a user will (and should) see from a storage
system, the client tools are an important part of the ARC
Storage. Currently ARC supports two ways of accessing the
storage solution:
\begin{itemize}
\item The \textbf{Command-line Interface} (CLI) provides
  access to the storage through the methods \texttt{stat}, \texttt{makeCollection},
  \texttt{unmakeCollection}, \texttt{putFile}, \texttt{getFile}, \texttt{delFile},
  \texttt{list} and \texttt{move}. Methods for modifying
  access and ownership will be available in the near future. The CLI
  assumes a relatively high level of computer competence
  from the user. However, the CLI, being a stand-alone tool, can be
  used to access ARC Storage from any computer (also including
  Microsoft Windows PCs) that has network access and
  a Python installation.
\item The \textbf{FUSE module} provides a high-level access to the
  ARC Storage. Filesystem in Userspace (FUSE) \cite{FUSE} provides a simple
  library and a kernel-userspace interface. Using FUSE and the ARC
  Storage Python interface, the FUSE module allows users to mount the
  ARC Storage namespace into the local namespace. This way, the user
  can use her/his favorite file manager to access her/his files and
  collections. The FUSE module provides most of the features provided
  by the CLI, with the exception of modifying some non-posix metadata.
\end{itemize}
It is worth mentioning that the client tools access the storage system
through the Bartender only. Currently upload and download is realized
through HTTP(S), but there are plans to add support for other
protocols, such as SRM and GridFTP.

@@

As is the case for all openly accessible web services, the security
model is of crucial importance for the ARC Storage. The security
architecture of the storage can be split into three parts; the
inter-service authorization; the transfer-level authorization; and the
high-level authorization.
\begin{itemize}
\item The \textbf{inter-service authorization} maintains the integrity of the
  internal communication between services. There is a lot of
  communication between the services of the ARC Storage. The
  Bartenders send requests to the Librarians and the Shepherds, the
  Shepherds communicate with the Librarians and the Librarians talk with
  the A-Hash. If any of these services is compromised or a new
  rogue service gets inserted in the system, the security of the
  entire system is compromised. To enable trust between the services,
  they need to know each other's Distinguished Names (DNs). This way,
  a rogue service would need to obtain a certificate with that exact
  DN from some trusted Certificate Authority (CA).
\item The \textbf{transfer-level authorization} handles the
  authorization in the cases of uploading and downloading files. When
  a transfer is requested, the Shepherd will provide a one-time
  Transfer URL (TURL) to which the client can connect. In the
  current architecture, this TURL is world-accessible. This may not seem
  very secure at first. However, provided that the TURL has a very long,
  unguessable name, that it is transfered to the user in a secure way
  and that it can only be accessed once before it is deleted, the
  chance of being compromized is very low.
\item The \textbf{high-level authorization} considers the access
  policies for the files and collections in the system. These policies
  are stored in A-Hash, in the metadata of the corresponding file or
  collection, providing a fine-grained security in the system.
\end{itemize}
The communication with and within the storage system is realized
through HTTPS with standard X.509 authentication.

@@

Even though the ARC Storage is in a pre-prototype state, it is already
possible to deploy and use the system for testing purposes. To
properly run a proof-of-concept test, the resources need to be
geographically distributed. In our test scenarios we utilized
resources in three different countries.

@@

In the tests we used two client machines: One ASUS eee 901 on a wireless network in Uppsala, 
and one Dell PowerEdge 1425SC with 10Gb ethernet connection at USIT. All the tests were performed and worked as expected: All the files were both 
uploaded and downloaded with correct checksums and they all got the requested 
number of replicas within a reasonable time. However, some of
the tests deserve extra attention.

@@

A significant feature of the ARC Storage is its automatic self
healing. Test 4 addresses this feature by studying the effect of
taking one Shepherd out of the system, and later reinserting it. Table
\ref{table:self healing} shows the storage distribution in three
states: Before interrupting a Shepherd, after the USIT-2
Shepherd is interrupted and redistribution of replicas is finished, and
alive, when the Shepherd is restarted and the system again has stabilized. We
can see that all files are properly distributed between the remaining
Shepherds when one of them is disrupted and that the storage load evens out 
again when the Shepherd comes back online. When the
system discovers that there are more replicas than needed, the first
Shepherd noticing will set mark its replica as obsolete. Since the failing Shepherd didn't lose any replicas while
being down, redistribution of replicas is just a matter of deleting
obsolete replicas. We also see that the NIIF node actually got fewer
files when USIT-2 went offline. If two Shepherds simultaneously
start uploading and replicating a missing replica, there will be too many
replicas. Here, a big replica on the NIIF node has randomly been chosen as
obsolete. However, the total storage usage remains the same in all
three cases.

@@

%% Even though we have more work in front of us, we have shown that the
%% ARC Storage already is in a state where initial real-life tests can be
%% done. We have described a simple, yet robust architecture which we
%% believe will benefit communities in need of a lightweight,
%% yet distributed storage solution.
While the work in the above mentioned areas is still in progress, we have 
already shown that the ARC Storage already is in a state where initial 
real-life tests can be done. We have described a simple, yet strong 
architecture which we believe will benefit communities in need of a lightweight, 
yet distributed storage solution.

>>>>

\begin{document}
\pagestyle{headings} 
\title{NorduGrid Storage -- Overview and Design of a Service-Oriented Storage System}
%% \title{Service-Oriented ARC Storage; \\An Overview and Design}
%% \author{  Jon K. Nilsen$^{2,3}$ \and Salman Toor$^{1}$
%%   \and Zsombor Nagy$^{5}$ \and Bjarte Mohn$^{4}$  \\ $^{1}$Scientific
%%   Computing Division, IT Department, Uppsala University, Sweden
%%   \\ $^{2}$Center for Information Technology, University of Oslo, Norway
%%   \\ $^{3}$Department of Physics,  University of Oslo, Norway
%%   \\  $^{4}$Department of Physics and Astronomy, Uppsala University,
%%   Sweden \\ $^{5}$National Information Infrastructure Development
%%   Institute, Hungary}
\author{  Jon K. Nilsen$^{\dagger,\ddagger}$ \and Salman Toor$^{*}$
  \and Zsombor Nagy$^{\mathparagraph}$ \and Bjarte Mohn$^{\mathsection}$ 
 \\ $^{\dagger}$University of Oslo, Center for Information Technology,\\
P.~O.~Box~1059, Blindern, N-0316~Oslo, Norway
  \\ $^{\ddagger}$University of Oslo, Dept. of Physics, P.~O.~Box~1048,
Blindern, N-0316~Oslo, Norway
\\$^{*}$Dept. of Information Technology, Div. of Scientific Computing, Uppsala
University,\\ Box~337, SE-75105~Uppsala, Sweden\\
$^{\mathparagraph}$Institute of National Information and Infrastructure
Development\\ NIIF/HUNGARNET, Victor~Hugo~18-22, H-1132~Budapest, Hungary 
  \\  $^{\mathsection}$Dept. of Physics and Astronomy, Div. of Nuclear and
Particle Physics,\\ Uppsala University, Box~535, SE-75121~Uppsala, Sweden \\ \\
j.k.nilsen@usit.uio.no \hspace{5 mm}
  salman.toor@it.uu.se \hspace{5 mm} zsombor@niif.hu \\  \hspace{7 mm}
  bjarte.mohn@fysast.uu.se}

@@

\maketitle             
%% \begin{abstract}
%% The KnowARC project will deliver the next generation Advanced Resource
%% Connector Fall  2009. A major component of this
%% service-oriented Grid middleware will be its storage system, the
%% ARC Storage, of which a
%% prototype is due Spring 2009. We will in this paper present an
%% overview of the storage system, its design and architecture. We will
%% also present some first, proof-of-concept test results, deploying the
%% the storage with a geographically distribution among three different countries.
%% ~\\~\\~
%% \end{abstract}
\begin{abstract}
There is an ever increasing need to utilize geographically distributed
hardware resources, both in terms of CPU and in terms of storage. The
service-oriented architecture provides a natural framework for
managing these resources. The next generation Advanced Resource
Connector (ARC) is a service-oriented Grid solution that will provide
the middleware to represent distributed resources in one simple
framework. In this paper, we will present an overview of ARC's novel storage system -- the NorduGrid storage -- itself a set of services providing a self-healing,
flexible Grid storage solution. We will also present some first
proof-of-concept test results from a deployment of the storage system
distributed across three different countries. 
~\\~\\~
\end{abstract}
%-----------------------------------------------------------
\section{Introduction}
\label{Introduction}

@@

In the advent of the next generation of the Advanced Resource
Connector (ARC) Grid middleware~\cite{arc} (new release due during
fall 2009), we present the new NorduGrid storage~\cite{ARCStoragedesigndoc}. This distributed storage system is
designed to provide an easy to use, flexible and scalable system that
can offer native storage and at the same time provide access to
third-party solutions like dCache~\cite{dCache,DSSWithdCache} by using
the same, uniform interface. Being part of ARC, the storage system is
based on a service-oriented architecture, in which each major
component of the system runs as a separate service within the ARC
Hosting Environment Daemon (HED)~\cite{HEDdesigndoc}. The HED service
container gives the capability of flexible replacement of the
components as well as the possibility to introduce modifications in
the future. %% The HED service container gives
%% the capability of flexible replacement of the components as well as
%% the future modification/enhancement of the system.

This paper is organized as follows. After describing
related work in Section \ref{Related Work}, we give a bird's eye
view of the next-generation Advanced Resource Connector (ARC) in
Section \ref{The Advance Resource Connector}. An overview of the NorduGrid storage is given in Section \ref{The NorduGrid Storage}, while
the architecture of the storage system is elaborated in Section
\ref{Architecture of the NorduGrid Storage}. In Section \ref{Testing and
  Discussion} we give some early, proof-of-concept results, before 
concluding in Section \ref{Conclusion and Future Work}.

@@

a joint effort between Deutsches Elektronen Synchrotron (DESY) \cite{dCachesite}, Fermi National Accelerator Laboratory \cite{FermiLab}, the  Nordic Data

@@

The NorduGrid storage, being a light-weight and flexible

@@

high bandwidth environments. In contrast, the NorduGrid storage is optimized

@@

To deliver the non-trivial quality of services required by the Grid,
there are a number of services running inside the HED. For example, Grid
job execution and management is handled by the A-REX service 
\cite{AREXdesigndoc}, policy
decisions are taken by the Charon service, the ISIS service is
responsible for information indexing, batch job submission is
handled by the Sched service, etc. In the later sections our discussion
will focus on the architecture and the design of another major set
of services, i.e., the NorduGrid storage.

\section{The NorduGrid Storage}
\label{The NorduGrid Storage}

The NorduGrid storage consists of a set of SOAP based services residing
within HED. Together, the services provide a self-healing, reliable, robust,
scalable, resilient and consistent data storage system. Data is
managed in a hierarchical global namespace with files and
subcollections grouped into collections\footnote{A concept very
  similar to files and directories in most common file systems.}.  A
dedicated root collection serves as a reference point for accessing the namespace. The hierarchy can then be
referenced using Logical Names. The global namespace is accessed
in the same manner as in local filesystems.

\begin{figure}
  \begin{center}
    \includegraphics[width=0.8\columnwidth]{arc1-storage-services-with-gateway.pdf}
  \end{center}
  \caption{Schematic of the NorduGrid storage architecture. The figure shows
    the main services of the NorduGrid storage and the communication channels. The B stands for Bartender, the S for Shepherd, the L for Librarian, the G for Gateway and the A-H stands for A-Hash. The straight lines denote the communication between services.}
  \label{fig:arc1_storage_service}
\end{figure}

Being based on a service-oriented architecture, the NorduGrid storage consists of a set of services as shown in Fig.
\ref{fig:arc1_storage_service}. The services are as follows: The Bartender (B)
provides the high-level interface to the user; the Librarian (L)
handles the entire storage namespace, using the A-Hash (A-H) as a metadatabase; the
Gateway (G) provides access to third-party storage systems; and
the Shepherd (S) is the frontend for the physical storage node. 
See Section \ref{Architecture of the NorduGrid Storage} for a detailed
discussion of the different services.   
The services communicate with each other through the
Message Chain Components in HED. The communication channels are
depicted by straight lines in Fig. \ref{fig:arc1_storage_service}.

The system supports file transfer through several transfer protocols,
with client side tools that hide various technical
details such as protocol specification, port numbers
and so on. To provide fault-tolerance, the system implements
automatic file replication, where
the replicas of a file\footnote{In this paper, a file
denotes an entry in the global namespace, while replica denotes the
physical file stored on a storage node.} are always stored on separate
storage nodes\footnote{A
  storage node is a server with a Shepherd, a Shepherd backend
and some storage service.}. To ensure a resilient, self-healing system, the Shepherd
regularly sends heartbeats to one of the Librarians. If the Shepherd fails
to send a heartbeat, one of the Librarians will mark its replicas as offline, thus initiating
re-replication of the replicas between the other Shepherds.

In the default implementation, the services will provide a
full-featured and consistent data storage system using files as an
atomic unit. The scope of the NorduGrid storage in the forseeable
future is restricted to providing support for file-based data services.
%% there are currently no plans for intrinsic support for database
%% access.

The NorduGrid storage supports third-party storage services in two
ways. Using the Gateway service, files already stored in some
third-party storage can be accessed using the provided client tools. To make use of third-party services for storing new
files, with the replication and fault-tolerance offered by the NorduGrid storage, third-party storage elements can also be used as backends for the
Shepherd service.

External grid middleware components can
access the storage system using ARC data service
interfaces directly. ARC will also provide interface
components that communicate via standard protocols such as SRM, which
will provide a single access point to the whole system.

\section{Architecture of the NorduGrid Storage}
\label{Architecture of the NorduGrid Storage}

@@

Each object in the NorduGrid storage has a globally
unique ID. A collection contains files and other collections,
and each of these entries has a name unique within the collection very
much like entries in a standard directory on a local filesystem. Besides
files and collections, the storage system has a third type of entry
called mount-points, which are references to the third-party storages
within the global namespace.

Replicas in a distributed storage system can have different states;
they can be broken, deleted, partially uploaded, etc. In the
NorduGrid storage, all replicas have assigned a state, some of which are
`\textbf{alive}' (if the replica passed the checksum test, and the
Shepherd reports that the storage node is healthy), `\textbf{invalid}'
(if the replica has a wrong checksum, or the Shepherd claims it has no
such replica), `\textbf{offline}' (if the Shepherd is down) and `\textbf{creating}' (if the replica is in
the state of uploading).

@@

The Shepherd services run as front-ends on storage nodes. A Shepherd
service reports to a Librarian about the 
node's health state in terms of replicas. While the Bartender initiates
file transfers, the actual transfers go directly between the storage node
and the clients.

@@

A-Hash is a hash table for consistently storing data in property-value
pairs. All metadata about files, collections, mount point, Shepherd's
health status, and so forth, is stored in the A-Hash. As the A-Hash is
the service storing the entire state of the storage system, it is
absolutely crucial for the sttorage system that the A-Hash is
consistent. The distribution and replication of this service is
therefore both necessary and challenging.

\begin{figure}
\centering
\includegraphics[width=1.0\columnwidth]{arc1-storage-downloading.pdf}
\caption{Schematic showing the scenario of file download from the NorduGrid storage.}
\label{arc1-storage-downloading}
\end{figure}

\subsection{Features}
\subsubsection{Heartbeat monitoring}
\label{Heartbeat and Replication}
In the proposed architecture, each Shepherd periodically sends
heartbeats to a Librarian with information about replicas
whose state changed since the last heartbeat, and the corresponding
GUIDs. These heartbeats are then stored in the A-Hash, making them
visible to all the Librarians in the system. If any of the Librarians
notices that a Shepherd is late with its heartbeat, it will mark
all the replicas in that Shepherd as offline.

@@

\subsubsection{Deletion}
\label{Deletion}
In a replicating storage system, there are several possible solutions
for file deletion, since both the replicas and the metadata need to be
removed. In our solution, we use the process of \textit{lazy
  deletion} \cite{LazyDeletion}. When a client requests that a file
should be deleted (and the user has the proper permissions) the
Bartender instructs the Librarian to remove the file's GUID from the
A-Hash. When the Shepherd, periodically checking all its replicas,
discovers that a replica has no file (and hence, no Location), it will
automatically delete the replica.

@@

Due to the unpredictable nature of the Grid environment, it is essential to
have some degree of  automatic recovery system in case of unexpected
failures. Fault tolerant behavior is required both at the level of
metadata and on the level of physical storage. While the work on fault-tolerant metadata is still in
progress, the following two
scenarios will explain the currently available recovery mechanism for
physical storage: 
\begin{itemize}
\item In the case of a file having an invalid checksum, the Shepherd
immediately informs the Librarian and the Librarian changes the state of the given replica to
`\textbf{invalid}'. To recover its replica, the Shepherd contacts a
Bartender and asks for another replica of the file. The Bartender chooses a valid
replica, initiates a file transfer from a Shepherd having the replica,
and returns the TURL to the Shepherd with the invalid replica. When the
Shepherd has received the replica and compared the checksum, it notifies the
Librarian that the replica is alive again.  
\item In the case of a Shepherd going offline, a Librarian will,
  as mentioned earlier, notice the lack of heartbeats and invalidate all the
  replicas, initiating new replication for all the files stored in
  this storage node. However, if the Shepherd
  again comes online, there will evidently be more replicas than
  needed. The first Shepherd to notice this will set its replica's
  state to `\textbf{thirdwheel}', i.e. obsolete. At the next occasion,
  the Shepherd will remove the replica, if and only if it has the only
  `\textbf{thirdwheel}' replica of this file. If there are more
  replicas with this state, all replicas will be set back to
  `\textbf{alive}' and the process is repeated. This scenario will be discussed further in Section
  \ref{Testing and Discussion}.
\end{itemize}
\subsubsection{Client tools }
\label{Client extension using FUSE module}

Being the only part a user will (and should) see from a storage
system, the client tools are an important part of the NorduGrid storage. Currently ARC supports two ways of accessing the
storage solution:
\begin{itemize}
\item The \textbf{Command-line Interface} (CLI) provides
  access to the storage through the methods \texttt{stat}, \texttt{makeCollection},
  \texttt{unmakeCollection}, \texttt{putFile}, \texttt{getFile}, \texttt{delFile},
  \texttt{list} and \texttt{move}. Methods for modifying
  access and ownership will be available in the near future. The CLI
  assumes a relatively high level of computer competence
  from the user. However, the CLI, being a stand-alone tool, can be
  used to access the storage system from any computer (also including
  Microsoft Windows PCs) that has network access and
  a Python installation.
\item The \textbf{FUSE module} provides a high-level access to the
  storage system. Filesystem in Userspace (FUSE) \cite{FUSE} provides a simple
  library and a kernel-userspace interface. Using FUSE and the ARC
  Python interface, the FUSE module allows users to mount the
  storage namespace into the local namespace. This way, the user
  can use her/his favorite file manager to access her/his files and
  collections. The FUSE module provides most of the features provided
  by the CLI, with the exception of modifying some non-posix metadata.
\end{itemize}
It is worth mentioning that the client tools queries the storage system
through the Bartender only. Currently upload and download is realized
through HTTP(S), but there are plans to add support for other
protocols, such as SRM and GridFTP.

@@

As is the case for all openly accessible web services, the security
model is of crucial importance for the NorduGrid storage. The security
architecture of the storage can be split into three parts; the
inter-service authorization; the transfer-level authorization; and the
high-level authorization.
\begin{itemize}
\item The \textbf{inter-service authorization} maintains the integrity of the
  internal communication between services. There is a lot of
  communication between the services of the storage system. The
  Bartenders send requests to the Librarians and the Shepherds, the
  Shepherds communicate with the Librarians and the Librarians talk with
  the A-Hash. If any of these services is compromised or a new
  rogue service gets inserted in the system, the security of the
  entire system is compromised. To enable trust between the services,
  they need to know each other's Distinguished Names (DNs). This way,
  a rogue service would need to obtain a certificate with that exact
  DN from some trusted Certificate Authority (CA).
\item The \textbf{transfer-level authorization} handles the
  authorization in the cases of uploading and downloading files. When
  a transfer is requested, the Shepherd will provide a one-time
  Transfer URL (TURL) to which the client can connect. In the
  current architecture, this TURL is world-accessible. This may not seem
  very secure at first. However, provided that the TURL has a very long,
  unguessable name, that it is transfered to the user in a secure way
  and that it can only be accessed once before it is deleted, the
  chance of being compromized is very low.
\item The \textbf{high-level authorization} considers the access
  policies for the files and collections in the system. These policies
  are stored in A-Hash, in the metadata of the corresponding file or
  collection, providing a fine-grained security in the system.
\end{itemize}
The communication with and within the storage system is realized
through HTTPS with standard X.509 authentication.

@@

Even though the NorduGrid storage is in a pre-prototype state, it is already
possible to deploy and use the system for testing purposes. To
properly run a proof-of-concept test, the resources need to be
geographically distributed. In our test scenarios we utilized
resources in three different countries.

@@

In the tests we used two client machines: One ASUS Eee 901 on a wireless network in Uppsala, 
and one Dell PowerEdge 1425SC with 10Gb ethernet connection at USIT. All the tests were performed and worked as expected: All the files were both 
uploaded and downloaded with correct checksums and they all got the requested 
number of replicas within a reasonable time. However, some of
the tests deserve extra attention.

@@

A significant feature of the NorduGrid storage is its automatic self-healing. Test 4 addresses this feature by studying the effect of
taking one Shepherd out of the system, and later reinserting it. Table
\ref{table:self healing} shows the storage distribution in three
states: Before interrupting a Shepherd, after the USIT-2
Shepherd is interrupted and redistribution of replicas is finished, and
alive, when the Shepherd is restarted and the system again has stabilized. We
can see that all files are properly distributed between the remaining
Shepherds when one of them is disrupted and that the storage load evens out 
again when the Shepherd comes back online. When the
system discovers that there are more replicas than needed, the first
Shepherd noticing will set mark its replica as obsolete. Since the failing Shepherd didn't lose any replicas while
being down, redistribution of replicas is just a matter of deleting
obsolete replicas. We also see that the NIIF node actually got fewer
files when USIT-2 went offline. If two Shepherds simultaneously
start uploading and replicating a missing replica, there will be too many
replicas. Here, a big replica on the NIIF node has randomly been chosen as
obsolete. However, the total storage usage remains the same in all
three cases.

@@

%% Even though we have more work in front of us, we have shown that the
%% ARC Storage already is in a state where initial real-life tests can be
%% done. We have described a simple, yet robust architecture which we
%% believe will benefit communities in need of a lightweight,
%% yet distributed storage solution.
While the work in the above mentioned areas is still in progress, we have 
already shown that the NorduGrid storage already is in a state where initial 
real-life tests can be done. We have described a simple, yet strong 
architecture which we believe will benefit communities in need of a lightweight, 
yet distributed storage solution.
